A thermodynamic view of dusty protoplanetary disks: axisymmetric stability

### 2016-08-19 paper.tex
minor changes

It is possible to re-formulate the two-fluid equations into that for a
single fluid of for the total density $\rho$ and center-of-mass
velocity $\bm{v}$ of the dusty gas mixture
\citep{laibe14}. The standard hydrodynamic equations for
$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
the dust-to-gas ratio $\rhod/\rhog$ and the dust-gas velocity
difference $\Delta\bm{v}$. The latter also contributes to
additional source terms in the momentum and energy equations.  
The one-fluid approach is appropriate for 
simulating gas tightly coupled to dust particles with small $\tstop$.  
This can be numerically challenging for the two-fluid or Lagragian
particles approach; but is straight forward to model in the
one-fluid framework \citep{price15}.

>>>>

It is possible to re-formulate the two-fluid equations into that for a
single fluid of for the total density $\rho$ and center-of-mass
velocity $\bm{v}$ of the dusty gas mixture
\citep{laibe14}. The standard hydrodynamic equations for
$\rho$ and $\bm{v}$ are supplemented by evolutionary equations for
the dust-to-gas ratio $\rhod/\rhog$ and the relative dust-gas velocity
$\Delta\bm{v}\equiv \bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$. The latter also contributes to 
additional source terms in the momentum and energy equations.  
The one-fluid approach is appropriate for 
simulating gas tightly coupled to dust particles with small $\tstop$.  
This can be numerically challenging for the two-fluid or Lagragian
particles approach; but is straight forward to model in the
one-fluid framework \citep{price15}.

### 2016-08-19 summary.tex
minor changes

>>>>

As described in \S\ref{dust_work}, a phase lag between the   
pressure and density response in a dusty gas is the
physical cause of dust-drag instabilities. %Whether the phase
%difference is positive or negative depends on the cooling function
%$\mathcal{C}$, but we may expect a system to generally have both
%modes, and that with a phase lag are unstable. 
This may be used to re-interpret the streaming instability or to
understand other dusty instabilities such as that recently discovered
by \cite{loren15,loren16}. These authors find an axisymmetric
instability as dust settles to the disk midplane.

%give numerical example

%expect the thermodynamic interpretation to hold even when tstop not
%small 
In fact, we expect a phase difference between pressure and the total
density  whenever $\tstop\neq0$; not just $\tstop\OmK\ll 1$ as 
considered in this work. This is simply because of the finite
relaxation time required for dust to respond to gas, and vice versa. 
An effective energy equation in the form of Eq. \ref{eff_energy} can
still be obtained for general $\tstop$. The
heat flux $\bm{F}$ in the cooling function $\mathcal{C}$ is determined
through solution to the relative velocity $\Delta \bm{v} =
\bm{v}_\mathrm{d}-\bm{v}_\mathrm{g}$ \citep{youdin05,laibe14}.  We
thus expect the `pressure-density lag' interpretation of dusty
instabilities to be generally applicable.
